---
title: 'Final Project'
subtitle: "PSTAT122: Design and Analysis of Experiments" 

author:
  - name: "Winter 2026"
affiliation-title: "Quarter"
format: 
 pdf:
    code-fold: true
    code-line-numbers: true
    code-copy: true
    code-tools: true
    self-contained: true
    toc: false
    toc-location: left
    number-sections: true
    geometry: margin=0.2in

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message =  FALSE)
knitr::opts_chunk$set(warning =  FALSE)
knitr::opts_chunk$set(error =  FALSE)

```



:::callout
<div style="text-align: center">
<span style="color: blue;"> **STUDENT NAME **</span> 
</div>
- Ethan Anghelache    (NetID 1)
- Miriam Montes    (NetID 2)
- Joanna Sells    (NetID 3)
- Saarah Solachuddin    (A0P1751)
:::


::: {.callout-caution}
# Due Date
**Due Date:** Monday, March 16, 2026, 11:59 PM
:::

```{r Libraries}
library(car)
library(emmeans)
library(tidyverse)
library(knitr)
```

# Introduction
-	Clear statement of the objective or research question.
-	Brief context or motivation.

# Experimental Design
- Description of factors and treatment structure.
- Clearly state what you are measuring and the units.
Examples: Number of words recalled (count), reaction time (seconds), taste rating (1–5 scale).
- Identify which factors are fixed vs. random.
- Description of design type (CRD, RCBD, factorial, etc.).
- Explain how randomization, replication, and (if used) blocking were implemented.
- Sample size: Provide number of observations per condition.
Guideline: 5–10 per treatment for CRD, 3–5 blocks for RCBD, total feasible within 1 hour.

(You are encouraged to explore more resources for determining the sample size )

```{r}
# randomization assignments
set.seed(21)

text_settings <- c("black, small", "black, large", "red, small", "red, large")

treatments <- rep(text_settings, each = 4)

sample(treatments)
```


# Data Collection
- **Procedure**: Describe how and when the experiment was conducted (e.g., location, date, steps taken).
- **Challenges/Adjustments**: Mention any difficulties or changes made during data collection (e.g., technical issues, time adjustments).
- **Data Presentation**: Display the collected data in tables or graphs, summarizing key measures like mean and standard deviation.

```{r}
data <- read.csv("Final_Data.csv")

# get factors in their own columns
# font color levels: black (-), red (+)
# font size levels: small (-), large (+)

read_speed <- data |>
  mutate("Font Color" = str_detect(Text.Settings, "red, small") | str_detect(Text.Settings, "red, large"),
         "Font Size" = str_detect(Text.Settings, "black, large") | str_detect(Text.Settings, "red, large")) |>
  rename("Time" = Time..s.) |>
  select(`Font Color`, `Font Size`, `Time`)

read_speed$`Font Color` <- factor(read_speed$`Font Color`,
                                  levels = c(TRUE, FALSE),
                                  labels = c("Red", "Black"))

read_speed$`Font Size` <- factor(read_speed$`Font Size`,
                                 levels = c(TRUE, FALSE),
                                 labels = c("Large", "Small"))

read_speed
kable(read_speed)
```


# Analysis
- **Exploratory Data**: Start with basic statistics (mean, SD) and visualizations (e.g., boxplots) to understand the data.

- **Hypothesis Testing**: Test your hypothesis with an appropriate statistical test (e.g., ANOVA).

- **Tables, Figures, & Code**: Include key results (ANOVA table, post-test) and relevant `R` code excerpts where needed.

-	Use `R` to analyze the data.


# Conclusions
-	Summarize key findings.
- Comment on limitations and possible improvements.


# References
 (If needed.)


# Appendices
 
 - `R` code.
